---
output: github_document
editor_options: 
  chunk_output_type: console
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "README-"
)
```

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.1248912.svg)](https://doi.org/10.5281/zenodo.1248912) 
[![Travis-CI Build Status](https://travis-ci.org/hrbrmstr/sergeant-caffeinated.svg?branch=master)](https://travis-ci.org/hrbrmstr/sergeant-caffeinated) 
[![Coverage Status](https://codecov.io/gh/hrbrmstr/sergeant-caffeinated/branch/master/graph/badge.svg)](https://codecov.io/gh/hrbrmstr/sergeant-caffeinated)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/sergeant-caffeinated)](https://cran.r-project.org/package=sergeant-caffeinated)

# 💂☕️ sergeant.caffeinated

Tools to Transform and Query Data with 'Apache' 'Drill' (JDBC)

## NOTE

This is the Java/JDBC-interface to Apache Drill. For non-Java/JDBC, see the `sergeant` package ([GitLab](https://gitlab.com/hrbrmstr/sergeant/); [GitHub](https://github.com/hrbrmstr/sergeant/)).

## Description

Drill + `sergeant` is (IMO) a streamlined alternative to Spark + `sparklyr` if you don't need the ML components of Spark (i.e. just need to query "big data" sources, need to interface with parquet, need to combine disparate data source types — json, csv, parquet, rdbms - for aggregation, etc). Drill also has support for spatial queries.

Using Drill SQL queries that reference parquet files on a local linux or macOS workstation can often be more performant than doing the same data ingestion & wrangling work with R (especially for large or disperate data sets). Drill can often help further streaming workflows that infolve wrangling many tiny JSON files on a daily basis.

Drill can be obtained from <https://drill.apache.org/download/> (use "Direct File Download"). Drill can also be installed via [Docker](https://drill.apache.org/docs/running-drill-on-docker/). For local installs on Unix-like systems, a common/suggestion location for the Drill directory is  `/usr/local/drill` as the install directory. 

Drill embedded (started using the `$DRILL_BASE_DIR/bin/drill-embedded` script) is a super-easy way to get started playing with Drill on a single workstation and most of many workflows can "get by" using Drill this way.

The following functions are implemented:

**`DBI`** (RJDBC)

- `drill_jdbc`:	Connect to Drill using JDBC, enabling use of said idioms. See `RJDBC` for more info.

NOTE: The DRILL JDBC driver fully-qualified path must be placed in the `DRILL_JDBC_JAR` environment variable. This is best done via `~/.Renviron` for interactive work. i.e. `DRILL_JDBC_JAR=/usr/local/drill/jars/drill-jdbc-all-1.14.0.jar`

**`dplyr`**: (RJDBC)

- `src_drill_jdbc`: Connect to Drill (using dplyr & RJDBC) + supporting functions

## Installation

```{r eval=FALSE}
devtools::install_git("https://gitlab.com/hrbrmstr/sergeant-caffeinated")
# OF
devtools::install_github("hrbrmstr/sergeant-caffeinated")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
options(width=120)
```

## Usage

```{r dplyr-01, message=FALSE}
library(sergeant.caffeinated)
library(tidyverse)

# use localhost if running standalone on same system otherwise the host or IP of your Drill server
ds <- src_drill_jdbc("localhost")  #ds
db <- tbl(ds, "cp.`employee.json`") 

# without `collect()`:
count(db, gender, marital_status)

count(db, gender, marital_status) %>% collect()

group_by(db, position_title) %>% 
  count(gender) -> tmp2

group_by(db, position_title) %>% 
  count(gender) %>% 
  ungroup() %>% 
  mutate(full_desc=ifelse(gender=="F", "Female", "Male")) %>% 
  collect() %>% 
  select(Title=position_title, Gender=full_desc, Count=n)

arrange(db, desc(employee_id)) %>% print(n=20)

mutate(db, position_title=tolower(position_title)) %>%
  mutate(salary=as.numeric(salary)) %>% 
  mutate(gender=ifelse(gender=="F", "Female", "Male")) %>%
  mutate(marital_status=ifelse(marital_status=="S", "Single", "Married")) %>% 
  group_by(supervisor_id) %>% 
  summarise(underlings_count=n()) %>% 
  collect()
```

```
### Test Results

```{r}
library(sergeant.caffeinated)
library(testthat)

date()

devtools::test()
```

## sergeant Metrics

```{r echo=FALSE}
cloc::cloc_pkg_md()
```

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). 
By participating in this project you agree to abide by its terms.
